---
title: "R practice BAN401"
author: "Marcus Lauritsen"
date: "26 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 2 &mdash; L5&6

In this problem, we are going to utilize while and for loops in order to manipulate some data.

1) Start by creating a vector of length 10 with random numbers between 0 and 20
a) Use a while loop and with an inner for loop with conditional statements to increment all numbers in the vector lower than 10 by 1, and decrement all numbers above 10 by 1 until the vector only contains numbers exactly equal to 10. Print the vector for each iteration of the while loop.
b) Use a while loop and vector conditioning to achieve the same result as in a)

```{r message=FALSE, warning=FALSE}
library(tidyverse)

v <- sample.int(20, 10, replace = TRUE)

while (all(v == 10) == FALSE) {
  v <- map_at(v, ~. + 1, .at = which(v < 10))
  v <- map_at(v, ~. - 1, .at = which(v > 10))
  print(flatten_dbl(v))
}
```

```{r}
v <- sample.int(20, 10, replace = TRUE)

while (all(v == 10) == FALSE) {
  v[which(v > 10)] <- vapply(v[which(v > 10)], function(x) x - 1, vector("double", 1))
  v[which(v < 10)] <- vapply(v[which(v < 10)], function(x) x + 1, vector("double", 1))
  print(v)
}
```

```{r}
v <- sample(20, 10, replace = TRUE)

while (!all(v == 10)) {
  v[which(v > 10)] <- v[which(v > 10)] - 1
  v[which(v < 10)] <- v[which(v < 10)] + 1
  print(v)
}
```

## Problem 3 &mdash; L1&2

Assume the following exchange rate: 1 USD = 9 NOK, and no taxes. Next year, the economy is expecting a growth in inflation equal to 3%, affecting both input and output prices. The company's market share is expected to grow from 5% to 25%, but the total market is constant. XYZ's costs are expected to decline by 5% of today’s cost because of a more efficient production line. Based on the above information, you are asked to answer the following questions:

What is the highest, lowest and the average price?
What is this year’s profits? Provide your answer with two strings, one telling what the profits are in NOK, and one in USD. Please round your answer to the nearest whole number if necessary.
```{r}
m <- matrix(c(55, 25, 250,
              63, 15, 133,
              85, 25, 821),
            nrow = 3,
            byrow = TRUE,
            dimnames = list(
              c("x", "y", "z"),
              c("price", "cost", "units.sold")
              ))
```


## Problem 3 &mdash; L3&4

Assume no taxes and that prices and costs are given in NOK. Next year, the economy is expecting a growth in inflation equal to 3%, affecting both input and output prices. XYZ's market share is expected to grow from 5% to 25%, but the total market is constant. XYZ's costs are expected to decline by 5% of today’s cost because of a more efficient production line. Based on this information, you are asked to answer the following questions:  

1. Find next year’s prices, costs, and number of units sold and display them in a matrix.
2. Assuming that the price and cost trends continue, but XYZ's market share stays the same for year +2, what is that year’s prices, costs and number of units? Use a matrix to show your answer.
3. Store the information on prices, costs, and number of units sold from the three time periods in a list. Print the list.
4. Display the development in the cost of product Y as a vector, using list indexing.

```{r problem 3L3&4}
m <- matrix(c(55, 25, 250,
              63, 15, 133,
              85, 25, 821),
            nrow = 3,
            byrow = TRUE,
            dimnames = list(
              c("x", "y", "z"),
              c("price", "cost", "units.sold")
              ))

forecast.products <- function(df, inflation, growth.factor, cost.prod, n = 1) {
  df[, "price"] <- df[, "price"] * (1 + inflation)^n
  df[, "cost"] <- df[, "cost"] * (1 + inflation - cost.prod)^n
  df[, "units.sold"] <- df[, "cost"] * growth.factor
  return(df)
}

# task 1
n <- forecast.products(df = m,
                       inflation = 0.03,
                       growth.factor = 5,
                       cost.prod = -0.05)

# task 2
n2 <- forecast.products(df = n,
                       inflation = 0.03,
                       growth.factor = 1,
                       cost.prod = -0.05)

# task 3
(matrices <- list(m, n, n2))

# task 4
costs.y <- vector("double", length(matrices))
for (i in seq_along(matrices)) costs.y[i] <- matrices[[i]]["y", "cost"]
print(costs.y)
```

## Problem #3 &mdash; L7&8
What is the most common hair color for men and/or women in the R dataset HairEyeColor?

Make a function that takes gender as a string argument. The function should display the frequency of the different hair colors for the given gender, in a pie chart. When passing "Female" to the function, it should display a pie chart similar the one below. The title of the chart should adapt according to the argument and the most common color - if the most common hair color for gender supposedly was red, the title should automatically say so.

```{r problem 3L7&8, message=FALSE, warning=FALSE}
library(tidyverse)
library(datasets)

HairEyeColor <- as.tibble(HairEyeColor)

pie.chart.hair <- function(sex) {
  filtered.df <- HairEyeColor %>% 
    filter(Sex == sex) %>%
    group_by(Hair) %>% 
    summarise(frequency = sum(n)) %>% 
    ungroup() %>% 
    arrange(desc(frequency))
  
  # create vector of colors to fill pie chart with
  colors <- vector("character", nrow(filtered.df))
  for (i in 1:nrow(filtered.df)) {
    colors[i] <- filtered.df$Hair[i]
  }
  colors <- replace(colors, colors == "Blond", "Yellow")
  
  pie(filtered.df$frequency,
      labels = filtered.df$Hair,
      col = colors,
      main = paste("The most common hair color for the sex", tolower(sex), "is", tolower(filtered.df[which.max(filtered.df$frequency), "Hair"])))
}

pie.chart.hair("Female")
pie.chart.hair("Male")
```

